---
title: "p8105_hw6_zz2603"
author: "Ziyi Zhao"
date: "11/22/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(mgcv)
library(ggpubr)
library(Hmisc)

```

# Problem 1
```{r}
birthwt <- read_csv("./data/birthweight.csv")

## check for NA
nalst <- vector("list",dim(birthwt)[2])
nalst = map(birthwt,is.na) %>% map(sum)
unlist(nalst) %>% as.vector()
```

There is no missing values in the dataset.

```{r}
## transfer babysex, frace, malform, mrace into factor
cln_bthwt <- birthwt %>% 
  mutate(babysex = as.factor(babysex),
         babysex = fct_relevel(babysex,"1"),
         frace = as.factor(frace),
         frace = fct_relevel(frace,"1"),
         malform = as.factor(malform),
         malform = fct_relevel(malform,"0"),
         mrace = as.factor(mrace),
         mrace = fct_relevel(mrace,"1"))
```

I want to see how each variable is correlated with birth weight and then decide which variable to use in the model. 

```{r fig.height=18,fig.width=20}
babybwt <- cln_bthwt[[4]]

crt_corr <- function(x) {
  cln_bthwt %>% ggplot(aes(y=bwt,x=x))+
    geom_point(alpha=0.1)+
    stat_smooth(method = "lm")+
    stat_cor(method="pearson")
}

corr_plotlst <- vector("list",dim(cln_bthwt)[2])
corr_plotlst <- map(cln_bthwt,crt_corr)

ggarrange(plotlist = corr_plotlst,widths = c(4,5))

```

Based on scatterplots, I found out that variables bhead, blength, delwt, fincome, gaweeks, mheight, momage, ppbmi, ppwt, smoken, and wtgain are significantly correlated in the model with outcome of bwt.

I also want to see how every continuous variable is correlated with each other. We remove variable pnumsga and pnumlbw because there is only one values 0.

```{r}
cont_bthbwt <- cln_bthwt %>% 
  select(bhead, blength, bwt, delwt, fincome, gaweeks, menarche,
         mheight, momage, parity, ppbmi, ppwt, smoken, wtgain)

Hmisc::rcorr(as.matrix(cont_bthbwt)) ## see correlation and p-value

## see it visually
cor(cont_bthbwt) %>% corrplot::corrplot()

```

From the plot, we can find out several pairs of variables are highly correlated,such as pairs of variable bmi and weights, and pairs of head size and length.

Let's see hypothesized structure of model of birthweight. According to my brief review of literature and websites, I found that baby's sex, parental race, mother's pre-pregnancy bmi (or weight), and gestational age can affect birthweight.

Since it will take too much time and effort to do variable selection or random forest test, I will combine the results from basic correlation analysis and hypothesize structure. 

Proposed model: babysex, blength, gaweeks, mrace, and ppbmi.

```{r}
prop_df <- cln_bthwt %>% 
  select(bwt,babysex,blength,gaweeks,mrace,ppbmi) %>% 
  mutate(id = c(1:dim(cln_bthwt)[1]))
 
train_df <- sample_frac(prop_df,0.8)
test_df <- anti_join(prop_df, train_df, by="id")

linear_mod = lm(bwt~babysex+blength+gaweeks+mrace+ppbmi,data=train_df)

linear_pred = add_predictions(train_df,linear_mod) %>% pull(pred)
linear_resid = add_residuals(train_df,linear_mod) %>% pull(resid)

pred_resid = tibble(linear_pred,linear_resid)

ggplot(pred_resid,aes(y=linear_resid,x=linear_pred))+
  geom_point()+stat_smooth()

```

I may propose a wrong model.

```{r}

cv_df = crossv_mc(cln_bthwt,100,test=0.2) %>% 
  mutate(train=map(train,as_tibble),
         test=map(test,as_tibble)) %>% 
  mutate(prop_mod = map(train,
                        ~lm(bwt~babysex+blength+gaweeks+mrace+ppbmi,
                            data = .x)),
         first_mod = map(train,~lm(bwt~blength+gaweeks,data=.x)),
         second_mod = map(train,
                          ~lm(bwt~bhead+blength+babysex+bhead*blength+
                                 bhead*babysex+blength*babysex+
                                 bhead*blength*babysex,data=.x))) %>%
  mutate(rmse_prop = map2_dbl(prop_mod,test,~rmse(model=.x,data=.y)),
         rmse_first = map2_dbl(first_mod,test,~rmse(model=.x,data=.y)),
         rmse_second = map2_dbl(second_mod,test,~rmse(model=.x,data=.y)))

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model=fct_inorder(model)) %>% 
  ggplot(aes(x=model,y=rmse))+geom_violin()

```

I don't think I fit correct model. I should fit smooth model here, but it returned error.

# Problem 2
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

## create a plot
weather_df %>% ggplot(aes(y=tmax,x=tmin))+
  geom_point(alpha=0.5)+stat_smooth(method = "lm")

## fit model
lm(tmax~tmin,data=weather_df) %>% broom::tidy() %>% knitr::kable(digits = 3)

## create 5000 bootstrap samples
boot_strap = weather_df %>% bootstrap(n=5000)

## get estimate R-square
boot_strap %>% mutate(models = map(strap,~lm(tmax~tmin,data=.x)),
                      results = map(models,broom::glance))%>% 
  select(-strap,-models) %>% 
  unnest(results) %>% janitor::clean_names() %>% 
  pull(r_squared) %>% as.vector() %>% 
  quantile(probs=c(0.025, 0.975), na.rm=TRUE)

## make plot for estimate r-square
boot_strap %>% mutate(models = map(strap,~lm(tmax~tmin,data=.x)),
                      results = map(models,broom::glance))%>% 
  select(-strap,-models) %>% 
  unnest(results) %>% janitor::clean_names() %>% 
  ggplot(aes(x=r_squared))+geom_density()

```

From the density plot, r-squared values are skewed to left.

```{r}
## get log(beta_hat0*beta_hat1)
beta_df <- boot_strap %>% mutate(models = map(strap,~lm(tmax~tmin,data=.x)),
                      results = map(models,broom::tidy)) %>% 
  select(-strap,-models) %>% 
  unnest(results) %>% 
  select(term,estimate) %>% 
  pivot_wider(names_from = "term",
              values_from = "estimate")

beta0hat <- beta_df[[1]] %>% unlist()
beta1hat <- beta_df[[2]] %>% unlist()

betahat_df <- tibble(beta0hat,beta1hat,log_betahat = log(beta0hat*beta1hat))

pull(betahat_df,log_betahat) %>% as.vector() %>% 
  quantile(probs=c(0.025, 0.975), na.rm=TRUE)

betahat_df %>% 
  ggplot(aes(log_betahat))+geom_density()
```

From the plot of log(beta0*beta1), the shape of density plot is balanced and symmetrics. 





